Oasis or Mirage? \\ Trade Network Propagation of Deforestation

### 2024-01-03 paper/main.tex
write policy background

\usepackage{amssymb,amsmath,amsfonts,eurosym,geometry,ulem,caption,color,setspace,sectsty,comment,caption,pdflscape,subfigure,array,hyperref, here}

@@

To empirically discuss this network propagation, I examine two historical policies, the 2007 Russian log export restriction and the 2014 Myanmar log export ban, respectively, and measure their effects on log production and exports in other countries. These two policy interventions have several useful features that make them ideal natural experiments for studying spillover effects. First, Russia and Myanmar were major suppliers of logs until their exports were restricted. Russia developed the largest coniferous forest in the world, the Siberian "taiga", which accounted for half of the world's supply of softwood lumber. Myanmar logged half of its tropical rainforests and exported teak logs to the world, accounting for 10\% of the world's tropical log supply. Therefore, a policy of export restrictions in these two major supplying countries would cause major supply shocks to other countries. Second, for countries other than these two, these policies were unpredictable exogenous shocks. Thus, the characteristics of these policies allow us to examine how environmental protection policies in one country affect forestry in other countries.\\

@@

The remainder of this paper is organized as follows. Section 2 provides an overview of two trade restriction policies (Russia in 2007 and Myanmar in 2014) and explains how these policies are exogeneous shocks for other countries. Section 3 describes how I construct trade network flow data for forest products, and Section 4 presents a sketch of the network data to give an overall picture of the trade network. Based on some basic facts explained in Section 4, Section 5 presents a simple model for constructing hypotheses on how one country's policy affects other countries. Section 6 describes the identification strategy for empirically testing the theoretical hypothesis. Section 7 presents the empirical results and Section 8 discusses their interpretation. Finally, Section 9 presents the conclusions. Extensions of the theoretical model and robustness checks of the empirical results are in the appendix.\\

\section{Russia and Myanmar's Restriction Policy}

\section{Trade Flow Data}

>>>>

\usepackage{amssymb,amsmath,amsfonts,eurosym,geometry,ulem,caption,setspace,sectsty,comment,caption,pdflscape,subfigure,array,hyperref, here}

@@

To empirically discuss this network propagation, I examine two historical policies, the 2007 Russian log export restriction and the 2014 Myanmar log export ban, respectively, and measure their effects on log production and exports in other countries. These two policy interventions have several useful features that make them ideal natural experiments for studying spillover effects. First, Russia and Myanmar were major suppliers of logs until their exports were restricted. Russia developed the largest coniferous forest in the world, the Siberian ``taiga'', which accounted for half of the world's supply of softwood lumber. Myanmar logged half of its tropical rainforests and exported teak logs to the world, accounting for 10\% of the world's tropical log supply. Therefore, a policy of export restrictions in these two major supplying countries would cause major supply shocks to other countries. Second, for countries other than these two, these policies were unpredictable exogenous shocks. Thus, the characteristics of these policies allow us to examine how environmental protection policies in one country affect forestry in other countries.\\

@@

The remainder of this paper is organized as follows. Section 2 describes how I construct trade network flow data for forest products. Section 3 provides an overview of two trade restriction policies (Russia in 2007 and Myanmar in 2014) and explains how these policies are exogeneous shocks for other countries. Section 4 presents a sketch of the network data to give an overall picture of the trade network. Based on some basic facts explained in Section 4, Section 5 presents a simple model for constructing hypotheses on how one country's policy affects other countries. Section 6 describes the identification strategy for empirically testing the theoretical hypothesis. Section 7 presents the empirical results and Section 8 discusses their interpretation. Finally, Section 9 presents the conclusions. Extensions of the theoretical model and robustness checks of the empirical results are in the appendix.\\

\section{Trade Flow Data}
My empirical analysis uses the Food and Agriculture Organization of the United Nations database (hereafter FAOSTAT) as the main data set. FAOSTAT provides annual production and trade statistics for forest products. FAOSTAT is derived from two data sources. The first is annual trade flow data between countries. FAOSTAT provides annual data on the value and volume of imports and exports at the country level by product type, such as roundwood (coniferous roundwood, non-coniferous non-tropical roundwood, and non-coniferous tropical roundwood, respectively), saw wood, plywood, and pulp. The dataset covers 173 countries and regions from 1997 to 2018. For example, it provides information on the amount of non-coniferous tropical roundwood Myanmar exported to China in each year. The second is annual production and trade data at the country level. FAOSTAT provides annual country-level data on production, export value, export volume, import value, and import volume. This dataset covers annual data for 244 countries and regions from 1961 to 2022. For example, the dataset includes annual information on how much coniferous roundwood the Russian Federation developed in each year. These annual data on trade flows and production data are combined to build the main data set.\\

In addition to FAOSTAT forestry data, I employ two other data sources. First, I use intra-country bilateral distance data to estimate the impact of distance on trade flows. The distance data are borrowed from the GeoDist database provided by the Centre d'Etudes Prospectives et d'Informations Internationales (CEPII). I use bilateral distance between centers of gravity weighted by the population of each country. Second, we use annual GDP data from the World Bank database. I provide descriptive evidence of this dataset in the next section.\\

Before presenting empirical results, I define key terms used in empirical part of this study. The following parts frequently use the term ``roundwood'', which corresponds to ``industrial roundwood'' defined by FAOSTAT. The term ``roundwood'' refers to total volume or value of all roundwood, excluding wood fuels. Thus, ``roundwood'' contains sawlogs and veneer logs; pulpwood, round and split; and other industrial roundwood. In the dataset, roundwood are classified into three categories: coniferous roundwood, non-coniferous non-tropical roundwood, and non-coniferous tropical roundwood.

\section{Russia and Myanmar's Restriction Policy}
This paper examines how one country's environmental policies to protect forests, namely Russia's log export restriction in 2007 and Myanmar's log export ban in 2014, affected the other countries through trade networks. In this section, I briefly summarize these two policies and present qualitative evidence supporting my assumption that these two policies were exogeneous supply shocks to other countries.

\subsection{Russia's log export restriction in 2007}
Russia developed the largest coniferous forest in the world, the Siberian ``taiga'', which accounted for half of the world's supply of coniferous wood.  To cope with the deforestation and to promote domestic wood processing industry, Russian government imposed an export tax of 6.5\% in July 2007, rising to 20\% on July 2007 and 25\% in April 2008 on logs exported from Russia. This restriction on log exports was introduced due to the Russian domestic situation. Additionally, as METI points out, the new Russian Forest Code, which was base of these export tax policies, was founded in December 2006. This haste of policy interventions supports our assumption that the policy was exogeneous shock for the rest of the world. \\

Fig \ref{fig:shock_russia} shows significance of this Russian supply shock. Panel A represents time-series of import price, import quantity, and import value for coniferous roundwood. First, let us see Panel A. Each gray line represents the average price at which roundwood produced in each country is imported into the rest of the world. The blue line represents the average price at which roundwood produced in Russia is imported into countries other than Russia. In the same way, Panels B and C show the average import volume and total import value, respectively, of roundwood produced in each country and imported into the rest of the world.

Panel A shows that Russia's restriction on log exports in 2007 increases the average price of roundwood imports from Russia about 1.1 $\sim$ 1.4 times. This increase is consistent with the export tax rate. In addition to the effect on price, Panel B describes that the policy dramatically reduces the average quantity of roundwood imports from Russia by almost half. Panel C shows the policy leads to a rapid decrease in value of imports from Russia. In other words, the export restriction policy in 2007 has significant impact on the supply from Russia to the rest of the world market.

\begin{figure}[H]
    \centering
    \caption{Impacts of Russia's Export Ban on Coniferous Roundwood Imports}
    \subfigure[][Import Price]{\includegraphics*[bb=0 0 650 650, scale=0.2]{"~/OneDrive - The University of Tokyo/deforestation/fig/shock/Russia/Industrial_roundwood__coniferous__export_import__imp_price.png"}}
    \subfigure[][Import Quantity]{\includegraphics*[bb=0 0 650 650, scale=0.2]{"~/OneDrive - The University of Tokyo/deforestation/fig/shock/Russia/Industrial_roundwood__coniferous__export_import__imp_quant.png"}}
    \subfigure[][Import Total Value]{\includegraphics*[bb=0 0 650 650, scale=0.2]{"~/OneDrive - The University of Tokyo/deforestation/fig/shock/Russia/Industrial_roundwood__coniferous__export_import__imp_val.png"}}
    \caption*{\small{Each gray line shows annual average price (a), quantity (b), and total value (c) of coniferous roundwood imports from each country. Blue line corresponds average price, quantity, and total value of imports from Russian Federation. Vertical line shows 2007, the year of Russia's restriction of log exports.}}
    \label{fig:shock_russia}
\end{figure}

\subsection{Myanmar's log export ban in 2014}
Myanmar had been a great supplier of teak (non-coniferous and tropical wood) in the world until 2014. However, half of the tropical rain forest had been lost by 2014 due to the log production and agriculture. To stop this serious deforestation, Myanmar government decided to ban exports of logs in 2014. WTO describes this situation as ``A ban on exports of round logs has been in place since 1 April 2014, to preserve Myanmar's dwindling forest cover, in particular, teak forests. The export prohibition also applies to boule-cut logs, baulk-square timber, and confiscated timber.'' \\

We can consider this export ban as an exogenous supply shock for the rest of the world. That is because the prohibition was due to the situation of Myanmar's country, which is not related to the other countries' situation. Fig \ref{fig:shock_Myanmar} shows significance of the supply shock. Panel A represents time-series of import price, import quantity, and import value for non-coniferous and non-tropical roundwood. First, let us see Panel A. Each gray line represents the average price at which roundwood produced in each country is imported into the rest of the world. The blue line represents the average price at which roundwood produced in Myanmar is imported into countries other than Myanmar. In the same way, Panels B and C show the average import volume and total import value, respectively, of roundwood produced in each country and imported into the rest of the world.\\

Panel A shows that Myanmar's export ban in 2014 did not have much impact on the average price of roundwood imports from Myanmar. Panel B, on the other hand, describes that the policy dramatically reduces the average quantity of roundwood imports from Myanmar from the third largest supplier. According to Panel C, the decrease in import volume leads to a decrease in value of imports from Myanmar. In other words, the export ban policy has significant impact on the supply from Myanmar to the rest of the world market.

\begin{figure}[H] 
    \centering
    \caption{Impacts of Myanmar's Export Restriction on Non-coniferous Tropical Roundwood Imports}
    \subfigure[][Import Price]{\includegraphics*[bb=0 0 650 650, scale=0.2]{"~/OneDrive - The University of Tokyo/deforestation/fig/shock/Myanmar/Industrial_roundwood__non-coniferous_tropical__export_import__imp_price.png"}}
    \subfigure[][Import Quantity]{\includegraphics*[bb=0 0 650 650, scale=0.2]{"~/OneDrive - The University of Tokyo/deforestation/fig/shock/Myanmar/Industrial_roundwood__non-coniferous_tropical__export_import__imp_quant.png"}}
    \subfigure[][Import Total Value]{\includegraphics*[bb=0 0 650 650, scale=0.2]{"~/OneDrive - The University of Tokyo/deforestation/fig/shock/Myanmar/Industrial_roundwood__non-coniferous_tropical__export_import__imp_val.png"}}
    \caption*{\small{Each gray line shows annual average price (a), quantity (b), and total value (c) of non-coniferous tropical roundwood imports from each country. Blue line corresponds average price, quantity, and total value of imports from Myanmar. Vertical line shows 2014, the year of Myanmar's ban of log exports.}}
    \label{fig:shock_Myanmar}
\end{figure}
